Meta-analytic approaches and effect sizes to account for 'nuisance heterogeneity' in comparative physiology

### 2021-10-08 MS/ms.Rmd
Fixing sentences that need to be reworded

The key features of all effect sizes used in meta-analyses are that: 1) they are statistical parameters that have been placed on a common scale so that they can be compared across studies and 2) they have some associated measure of precision (i.e., sampling variance) [@Koricheva2013; @Borenstein2009; @Schmid2021]. Effect sizes in comparative physiology are quite diverse; they can be the mean difference between two groups [e.g., difference in mean between a control and treatment group; e.g., @WuSeebacher2020], the slope of a regression [e.g., between body size and metabolic rate, @Uyeda2017], a correlation between two variables [e.g., between hormone levels and behaviour, @Holtman2016], or even the raw mean (or variance) itself [e.g., comparing CTmax or CTmin across species - e.g., @Sunday2014]. In many cases, different effect sizes might be used depending on the question. For example, to investigate the effect of temperature on a physiological trait, one can compile studies experimentally manipulating temperature and analyse the mean 'effect' by creating contrast-based effect sizes, or alternatively, just model the trait in each temperature treatment. The choice as to which one to use boils down to the types of data available from primary studies, the specific question of interest, and whether studies are observational or experimental in nature. Considering our temperature example above, if we are interested in the magnitude of change between temperature treatments, we may want to use a contrast-based effect size, such as the log response ratio, *$\ln{RR}$*. Instead, if we are interested in how the mean of some physiological trait changes with temperature, we may model the mean itself. In the latter scenario, as long as all the means are measured in the same units (e.g., degrees Celsius) or can be converted to the same units, then the mean may provide the desired answer to the original question.

@@

```{r, figsurv, fig.wid = 16.833333, fig.height=5,fig.cap="**Summary of meta-analytic practices in comparative physiology.** (A) Different types of effect measures commonly used in meta-analyses and their relative frequency of use. (B) Frequency of studies where ‘nuisance heterogenity’ had the potential to impact effect sizes (x-axis), and the frequencies of studies explicitly dealing with nuisance heterogeneity (bar colour). (C) Frequencies of analytical approaches used to analyse effect measures. 'Orange' bars indicate weighted models."}
p1 = dat %>% dplyr::group_by(effect_size_type) %>% janitor::tabyl(effect_size_type)  %>% 
  dplyr::summarise(SMD = dplyr::if_else(grepl("SMD", effect_size_type), n, 0),
                  lnRR = dplyr::if_else(grepl("Response Ratio", effect_size_type), n, 0),
                    Zr = dplyr::if_else(grepl("correlation", effect_size_type), n, 0),
                 Other = dplyr::if_else(grepl("other|Q10", effect_size_type), n, 0),
                 Mean = dplyr::if_else(grepl("mean/proportion/raw", effect_size_type), n, 0),
                 lnCVR = dplyr::if_else(grepl("lnCVR", effect_size_type), n, 0),
                 Slope = dplyr::if_else(grepl("slope", effect_size_type), n, 0)) %>% 
  colSums(.) %>% data.frame(effect = names(.), n = .) %>% mutate(prop = n / sum(n)) %>% arrange(effect) %>% 
  mutate(effect = fct_relevel(effect, c("Other", "Zr", "lnRR", "SMD", "lnCVR", "Slope", "Mean"))) %>% 
  ggplot() + tm + 
  geom_col(aes(x = effect, y = prop), width = 0.5, fill = "black", colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,0.35)) +
  labs(x = "Effect Size",
       y = "Frequency") +
  coord_flip()

@@

As can be seen in Table \@ref(tab:tablea), none of the effect measures control for nuisance heterogeneity. For example, the magnitude of *lnRR* or *SMD* might depend on differences in the temperatures, dosages, or pH applied to each treatment. The same applies to modelling the means. For example, standard metabolic rate (SMR) is temperature-dependent. As such, one needs to standardise the temperature measurement to compare across studies [@White2006; @Uyeda2017] (if, of course, temperature is not of interest itself). While we could model this explicitly, this may not always be desirable. It can often be useful to standardise the effect to both ease its interpretation and simplify modelling. In Box 1, we provide readers with guidance on how comparative physiologists can apply corrections to the effect measure itself. We highlight two examples in Box 1 that can be easily generalized to alternative effect size measures and cases. Below, we show how we can formalise a traditional effect size, such as $lnRR$, to capture this nuisance variation. We use a common, well understood, effect measure of interest to comparative physiologists -- the temperature coefficient ($Q_{10}$). Figure \@ref(fig:fig0) summarises the different effect sizes that we mention in Table 1, Box 1 and the section below.

>>>>

The key features of all effect sizes used in meta-analyses are that: 1) they are statistical parameters that have been placed on a common scale so that they can be compared across studies and 2) they have some associated measure of precision (i.e., sampling variance) [@Koricheva2013; @Borenstein2009; @Schmid2021]. Effect sizes in comparative physiology are quite diverse; they can be the mean difference between two groups [e.g., difference in mean between a control and treatment group; e.g., @WuSeebacher2020], the slope of a regression [e.g., between body size and metabolic rate, @Uyeda2017], a correlation between two variables [e.g., between hormone levels and behaviour, @Holtman2016], or even the raw mean (or variance) itself [e.g., comparing CTmax or CTmin across species - e.g., @Sunday2014].

In many cases, different effect sizes might be used depending on the question. For example, to investigate the effect of temperature on a physiological trait, one can compile studies experimentally manipulating temperature and analyse the mean 'effect' by creating contrast-based effect sizes, or alternatively, just model the trait in each temperature treatment. The choice as to which one to use boils down to the types of data available from primary studies, the specific question of interest, and whether studies are observational or experimental in nature. Considering our temperature example above, if we are interested in the magnitude of change between temperature treatments, we may want to use a contrast-based effect size, such as the log response ratio, *$\ln{RR}$*. Instead, if we are interested in how the mean of some physiological trait changes with temperature, we may model the mean itself. In the latter scenario, as long as all the means are measured in the same units (e.g., degrees Celsius) or can be converted to the same units, then the mean may provide the desired answer to the original question.

@@

```{r, figsurv, fig.wid = 16.833333, fig.height=5,fig.cap="**Summary of meta-analytic practices in comparative physiology.** (A) Different types of effect measures commonly used in meta-analyses and their relative frequency of use. (B) Frequency of studies where nuisance heterogenity had the potential to impact effect sizes (x-axis), and the frequencies of studies explicitly dealing with nuisance heterogeneity (bar colour). (C) Frequencies of analytical approaches used to analyse effect measures. 'Orange' bars indicate weighted models."}
p1 = dat %>% dplyr::group_by(effect_size_type) %>% janitor::tabyl(effect_size_type)  %>% 
  dplyr::summarise(SMD = dplyr::if_else(grepl("SMD", effect_size_type), n, 0),
                  lnRR = dplyr::if_else(grepl("Response Ratio", effect_size_type), n, 0),
                    Zr = dplyr::if_else(grepl("correlation", effect_size_type), n, 0),
                 Other = dplyr::if_else(grepl("other|Q10", effect_size_type), n, 0),
                 Mean = dplyr::if_else(grepl("mean/proportion/raw", effect_size_type), n, 0),
                 lnCVR = dplyr::if_else(grepl("lnCVR", effect_size_type), n, 0),
                 Slope = dplyr::if_else(grepl("slope", effect_size_type), n, 0)) %>% 
  colSums(.) %>% data.frame(effect = names(.), n = .) %>% mutate(prop = n / sum(n)) %>% arrange(effect) %>% 
  mutate(effect = fct_relevel(effect, c("Other", "Zr", "lnRR", "SMD", "lnCVR", "Slope", "Mean"))) %>% 
  ggplot() + tm + 
  geom_col(aes(x = effect, y = prop), width = 0.5, fill = "black", colour = text.colour, size = 0.5) +
  scale_y_continuous(expand=c(0,0), limits = c(0,0.35)) +
  labs(x = "Effect Size",
       y = "Frequency") +
  coord_flip()

@@

As can be seen in Table \@ref(tab:tablea), none of the effect measures control for nuisance heterogeneity -- a point reinforced by our survey. For example, the magnitude of *lnRR* or *SMD* can depend on differences in the temperatures, dosages, or pH applied to each treatment. The same applies to modelling the means. For example, standard metabolic rate (SMR) is temperature-dependent. As such, one needs to standardise the temperature measurement to compare across studies [@White2006; @Uyeda2017] (if, of course, temperature is not of interest itself). While we could model nuisance heterogeneity explicitly, this may not always be desirable. It can often be useful to standardise the effect to both ease its interpretation and simplify modelling. In Box 1, we provide readers with guidance on how comparative physiologists can apply corrections to the effect measure itself. We highlight two examples in Box 1 that can be easily generalized to alternative effect size measures and show how they are distinct from traditional effect measures (Figure \@ref(fig:fig0)). Below we focus in on one familiar example to elaborate on how a traditional effect size, such as $lnRR$, can be reformalised to capture nuisance variation. We use a common, well understood, effect measure of interest to comparative physiologists -- the temperature coefficient ($Q_{10}$).
